---
title: "Analyze Science Grades 20-21"
author: Maura Carter
date: 07_20_21
output: html_notebook
---



```{r}
library(dplyr)
science_grades <- get_table(.table_name = "StudentHistoricalGrades", .database_name = "PROD1", .schema = "Schools")%>%
    filter(CreditType == "SC", AcademicYear == '2019-2020' | AcademicYear == '2020-2021') %>%
    collect()
head(science_grades)

```

Student Survey CP (20-21)
```{r}
survey_20_21_CP <- read_excel("C:/Users/maura.carter/ideapublicschools.org/Evaluation of Camp Rio - Documents/Camp RIO Survey - Video - College Prep.xlsx") %>%
  mutate(StudentID = as.numeric(`Please enter your student number.`))

survey_20_21_CP

#Clean the variable names (too many spaces/long!)
library(janitor)
survey2_20_21_CP <- survey_20_21_CP %>%
  clean_names()
head(survey2_20_21_CP)

survey3_20_21_CP <- survey2_20_21_CP%>%
  mutate(num_visits = before_today_how_many_times_have_you_visited_camp_rio, na.rm = TRUE)
survey3_20_21_CP

survey3_20_21_CP$num_visits <-as.numeric(survey3_20_21_CP$num_visits)
survey3_20_21_CP

#All a col called "num_visits_bin" to change number of visits to Binary (3+ = 1, 1 or 2 = 0;  OR:  >2 = 1, <3 = 0)
survey4_20_21_CP <- survey3_20_21_CP %>%
  mutate(num_visits_bin = num_visits, na.rm = TRUE)
survey4_20_21_CP

# "num_visits_bin" col: Make everything less than 2 (0, 1 prior visits) a "0". It really means that if they had 0 visits, it's 1 visit when you include today. If they had 1 prior visit, it's 2 including today, etc.
survey4_20_21_CP$num_visits_bin[survey4_20_21_CP$num_visits_bin < 2] <- 0
survey4_20_21_CP

# "num_visits_bin" col: Make everything larger than 0 (all visits that are 2, 3, 4+ for prior visits) to a "1". It will actually start at "3" because if you include today's visit, you do 2 + 1 = 3.
survey4_20_21_CP$num_visits_bin[survey4_20_21_CP$num_visits_bin > 0] <- 1
survey4_20_21_CP

survey5_20_21_CP <- survey4_20_21_CP %>%
  select(student_id, num_visits_bin)
survey5_20_21_CP

```

Student Survey Academy (20-21)
```{r}

survey_20_21_AC <- read_excel("C:/Users/maura.carter/ideapublicschools.org/Evaluation of Camp Rio - Documents/Camp RIO Survey - Video - Academy.xlsx")  %>% 
  mutate(StudentID = as.numeric(`Please enter your student number.`))

survey_20_21_AC

#Clean the variable names (too many spaces/long!)
library(janitor)
survey2_20_21_AC <- survey_20_21_AC %>%
  clean_names()
head(survey2_20_21_AC)

survey3_20_21_AC <- survey2_20_21_AC %>%
  mutate(num_visits = before_today_how_many_times_have_you_visited_camp_rio, na.rm = TRUE)
survey3_20_21_AC

survey3_20_21_AC$num_visits <-as.numeric(survey3_20_21_AC$num_visits)
survey3_20_21_AC

#All a col called "num_visits_bin" to change number of visits to Binary (3+ = 1, 1 or 2 = 0;  OR:  >2 = 1, <3 = 0)
survey4_20_21_AC <- survey3_20_21_AC %>%
  mutate(num_visits_bin = num_visits, na.rm = TRUE)
survey4_20_21_AC

# "num_visits_bin" col: Make everything less than 2 (0, 1 prior visits) a "0". It really means that if they had 0 visits, it's 1 visit when you include today. If they had 1 prior visit, it's 2 including today, etc.
survey4_20_21_AC$num_visits_bin[survey4_20_21_AC$num_visits_bin < 2] <- 0
survey4_20_21_AC

# "num_visits_bin" col: Make everything larger than 0 (all visits that are 2, 3, 4+ for prior visits) to a "1". It will actually start at "3" because if you include today's visit, you do 2 + 1 = 3.
survey4_20_21_AC$num_visits_bin[survey4_20_21_AC$num_visits_bin > 0] <- 1
survey4_20_21_AC

survey5_20_21_AC <- survey4_20_21_AC %>%
  select(student_id, num_visits_bin)
survey5_20_21_AC





```

Student Survey: College Prep 19-20
```{r}
survey_19_20_CP <- read_excel("C:/Users/maura.carter/ideapublicschools.org/Evaluation of Camp Rio - Documents/Camp RIO Survey - College Prep.xlsx")%>%
  mutate(StudentID = as.numeric(`Please enter your student number.`))

survey_19_20_CP

#Clean the variable names (too many spaces/long!)
library(janitor)
survey2_19_20_CP <- survey_19_20_CP %>%
  clean_names()
head(survey2_19_20_CP)

survey3_19_20_CP <- survey2_19_20_CP%>%
  mutate(num_visits = before_today_how_many_times_have_you_visited_camp_rio, na.rm = TRUE)
survey3_19_20_CP

survey3_19_20_CP$num_visits <-as.numeric(survey3_19_20_CP$num_visits)
survey3_19_20_CP

#All a col called "num_visits_bin" to change number of visits to Binary (3+ = 1, 1 or 2 = 0;  OR:  >2 = 1, <3 = 0)
survey4_19_20_CP <- survey3_19_20_CP %>%
  mutate(num_visits_bin = num_visits, na.rm = TRUE)
survey4_19_20_CP

# "num_visits_bin" col: Make everything less than 2 (0, 1 prior visits) a "0". It really means that if they had 0 visits, it's 1 visit when you include today. If they had 1 prior visit, it's 2 including today, etc.
survey4_19_20_CP$num_visits_bin[survey4_19_20_CP$num_visits_bin < 2] <- 0
survey4_19_20_CP

# "num_visits_bin" col: Make everything larger than 0 (all visits that are 2, 3, 4+ for prior visits) to a "1". It will actually start at "3" because if you include today's visit, you do 2 + 1 = 3.
survey4_19_20_CP$num_visits_bin[survey4_19_20_CP$num_visits_bin > 0] <- 1
survey4_19_20_CP

survey5_19_20_CP <- survey4_19_20_CP %>%
  select(student_id, num_visits_bin)
survey5_19_20_CP



```

Student Survey: Academy 19-20
```{r}
survey_19_20_AC <- read_excel("C:/Users/maura.carter/ideapublicschools.org/Evaluation of Camp Rio - Documents/Camp RIO Survey - Academy.xlsx")%>%
  mutate(StudentID = as.numeric(`Please enter your student number.`))

survey_19_20_AC

#Clean the variable names (too many spaces/long!)
library(janitor)
survey2_19_20_AC <- survey_19_20_AC %>%
  clean_names()
head(survey2_19_20_AC)

survey3_19_20_AC <- survey2_19_20_AC%>%
  mutate(num_visits = before_today_how_many_times_have_you_visited_camp_rio, na.rm = TRUE)
survey3_19_20_AC

survey3_19_20_AC$num_visits <-as.numeric(survey3_19_20_AC$num_visits)
survey3_19_20_AC

#All a col called "num_visits_bin" to change number of visits to Binary (3+ = 1, 1 or 2 = 0;  OR:  >2 = 1, <3 = 0)
survey4_19_20_AC <- survey3_19_20_AC %>%
  mutate(num_visits_bin = num_visits, na.rm = TRUE)
survey4_19_20_AC

# "num_visits_bin" col: Make everything less than 2 (0, 1 prior visits) a "0". It really means that if they had 0 visits, it's 1 visit when you include today. If they had 1 prior visit, it's 2 including today, etc.
survey4_19_20_AC$num_visits_bin[survey4_19_20_AC$num_visits_bin < 2] <- 0
survey4_19_20_AC

# "num_visits_bin" col: Make everything larger than 0 (all visits that are 2, 3, 4+ for prior visits) to a "1". It will actually start at "3" because if you include today's visit, you do 2 + 1 = 3.
survey4_19_20_AC$num_visits_bin[survey4_19_20_AC$num_visits_bin > 0] <- 1
survey4_19_20_AC

survey5_19_20_AC <- survey4_19_20_AC %>%
  select(student_id, num_visits_bin)
survey5_19_20_AC

```



Join survey5_20_21_CP (college prep) to survey5_20_21_AC (academy)
```{r}
survey_20_21 <- bind_rows(survey5_20_21_AC, survey5_20_21_CP, .id=NULL)
survey_20_21

survey2_20_21 <- bind_rows(survey5_19_20_AC, survey5_19_20_CP, .id=NULL)
survey2_20_21

#Bind 19-20 to 20-21 (student surveys)
survey3_19_20_21 <- bind_rows(survey_20_21, survey2_20_21, .id=NULL)
survey3_19_20_21

```


GPA data
```{r}
gpa <- get_table(.table_name = "StudentAcademicSummary", .database_name = "PROD1", .schema = "Schools") %>%
  filter(AcademicYear == '2020-2021' | AcademicYear == '2019-2020' ) %>%
  collect()
head(gpa)

gpa_20_21 <- gpa %>%
  select(PowerSchoolStudentID, AcademicYear, StudentNumber, GradeLevelID,CumulativeWeightedGPA, State)%>%
  filter(AcademicYear == "2020-2021")
gpa_20_21

gpa2_20_21 <- gpa_20_21 %>%
  mutate(student_id = StudentNumber)
gpa2_20_21

gpa_20_21_search <- gpa_20_21%>%
  filter(StudentNumber== "108028113")
gpa_20_21_search

gpa_0 <- gpa %>%
  mutate(student_id = StudentNumber)%>%
  select(PowerSchoolStudentID, AcademicYear, student_id, GradeLevelID,CumulativeWeightedGPA, State)
gpa_0

gpa_00 <- gpa_0 %>%
  select(student_id)
gpa_00


```

Join GPA data to Student Survey data table (20-21).
```{r}
#test. Use "inner_join". left_join doesn't work.
both <- left_join(gpa_00, survey3_19_20_21, by = NULL)
both

both2 <- inner_join(gpa_00, survey3_19_20_21, by = NULL)
both2

#both Academic Years 19-20, 20-21
gpa_survey_19_20_21 <- inner_join(survey3_19_20_21, gpa_0, by=NULL)
gpa_survey_19_20_21  #It worked!



```



9th Grade (GPA)
(20-21):     >2 visits:    4.019         *4.019-3.895 = 0.124  (20-21 is larger)
            (1,2) visits:  3.903         *3.903-3.718 = 0.185  (20-21 is larger)
            
(19-20):     >2 visits:    3.895
            (1,2) visits:  3.718

```{r}
#20-21
gpa9_survey_19_20_21 <- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "9")%>%
  filter(AcademicYear == "2020-2021")
gpa9_survey_19_20_21

gpa9_survey_19_20_21$num_visits_bin <-as.character(gpa9_survey_19_20_21$num_visits_bin)
gpa9_survey_19_20_21

#Regression
gpa_9 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa9_survey_19_20_21)
summary(gpa_9)

#Average GPA, 9th, (>2) visits to Camp RIO   (20-21)           GPA: 4.019
gpa9_survey_19_20_21_average_bin1 <- gpa9_survey_19_20_21%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/115)
gpa9_survey_19_20_21_average_bin1

#Average GPA, 9th, (1,2) visits to Camp RIO    (20-21)          GPA: 3.903
gpa9_survey_19_20_21_average_bin0 <- gpa9_survey_19_20_21%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/29)
gpa9_survey_19_20_21_average_bin0
#------------------------------------
#19-20
gpa9_survey_19<- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "9")%>%
  filter(AcademicYear == "2019-2020")
gpa9_survey_19

gpa9_survey_19$num_visits_bin <-as.character(gpa9_survey_19$num_visits_bin)
gpa9_survey_19

#Regression
gpa_9_19 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa9_survey_19)
summary(gpa_9_19)

#Average GPA, 9th, (>2) visits to Camp RIO  (19-20)            GPA: 3.895
gpa9_survey_19_average_bin1 <- gpa9_survey_19%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/183)
gpa9_survey_19_average_bin1

#Average GPA, 9th, (1,2) visits to Camp RIO  (19-20)             GPA: 3.718
gpa9_survey_19_average_bin0 <- gpa9_survey_19%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/17)
gpa9_survey_19_average_bin0

```

10th Grade (GPA)
(20-21):     >2 visits:    3.930         *3.930-3.934 = -0.004  (19-20 is larger)
            (1,2) visits:  3.756         *3.756-2.558 =  1.198  (20-21 is larger)
            
(19-20):     >2 visits:    3.934
            (1,2) visits:  2.558

```{r}
#20-21
gpa10_survey_19_20_21 <- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "10")%>%
  filter(AcademicYear == "2020-2021")
gpa10_survey_19_20_21

gpa10_survey_19_20_21$num_visits_bin <-as.character(gpa10_survey_19_20_21$num_visits_bin)
gpa10_survey_19_20_21

#Regression
gpa_10 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa10_survey_19_20_21)
summary(gpa_10)

#Average GPA, 10th, (>2) visits to Camp RIO   (20-21)           GPA: 3.930
gpa10_survey_19_20_21_average_bin1 <- gpa10_survey_19_20_21%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/183)
gpa10_survey_19_20_21_average_bin1

#Average GPA, 10th, (1,2) visits to Camp RIO    (20-21)          GPA: 3.756
gpa10_survey_19_20_21_average_bin0 <- gpa10_survey_19_20_21%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/20)
gpa10_survey_19_20_21_average_bin0
#------------------------------------
#19-20
gpa10_survey_19<- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "10")%>%
  filter(AcademicYear == "2019-2020")
gpa10_survey_19

gpa10_survey_19$num_visits_bin <-as.character(gpa10_survey_19$num_visits_bin)
gpa10_survey_19

#Regression
#gpa_10_19 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa10_survey_19)
#summary(gpa_10_19)

#Average GPA, 10th, (>2) visits to Camp RIO  (19-20)            GPA: 3.934
gpa10_survey_19_average_bin1 <- gpa10_survey_19%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/74)
gpa10_survey_19_average_bin1

#Average GPA, 10th, (1,2) visits to Camp RIO  (19-20)             GPA: 2.558
gpa10_survey_19_average_bin0 <- gpa10_survey_19%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/17)
gpa10_survey_19_average_bin0




```

11th Grade (GPA)
(20-21):     >2 visits:    3.973         *3.973-3.943 = 0.030   (20-21 is larger)
            (1,2) visits:  4.085         *4.085-3.923 = 0.162   (20-21 is larger)
            
(19-20):     >2 visits:    3.943
            (1,2) visits:  3.923
```{r}
#20-21
gpa11_survey_19_20_21 <- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "11")%>%
  filter(AcademicYear == "2020-2021")
gpa11_survey_19_20_21

gpa11_survey_19_20_21$num_visits_bin <-as.character(gpa11_survey_19_20_21$num_visits_bin)
gpa11_survey_19_20_21

#Regression
gpa_11 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa11_survey_19_20_21)
summary(gpa_11)

#Average GPA, 11th, (>2) visits to Camp RIO   (20-21)           GPA: 3.973
gpa11_survey_19_20_21_average_bin1 <- gpa11_survey_19_20_21%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/72)
gpa11_survey_19_20_21_average_bin1

#Average GPA, 11th, (1,2) visits to Camp RIO    (20-21)          GPA: 4.085
gpa11_survey_19_20_21_average_bin0 <- gpa11_survey_19_20_21%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/13)
gpa11_survey_19_20_21_average_bin0
#------------------------------------
#19-20
gpa11_survey_19<- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "11")%>%
  filter(AcademicYear == "2019-2020")
gpa11_survey_19

gpa11_survey_19$num_visits_bin <-as.character(gpa11_survey_19$num_visits_bin)
gpa11_survey_19

#Regression
#gpa_10_19 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa10_survey_19)
#summary(gpa_10_19)

#Average GPA, 11th, (>2) visits to Camp RIO  (19-20)            GPA: 3.943
gpa11_survey_19_average_bin1 <- gpa11_survey_19%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/142)
gpa11_survey_19_average_bin1

#Average GPA, 11th, (1,2) visits to Camp RIO  (19-20)             GPA: 3.923
gpa11_survey_19_average_bin0 <- gpa11_survey_19%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/36)
gpa11_survey_19_average_bin0



```

12th Grade (GPA)
11th Grade (GPA)
(20-21):     >2 visits:    3.943         *3.943-4.057 = -0.114   (19-20 is larger)
            (1,2) visits:  3.923         *3.923-3.880 =  0.043   (20-21 is larger)
            
(19-20):     >2 visits:    4.057
            (1,2) visits:  3.880
```{r}

#20-21
gpa12_survey_19_20_21 <- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "12")%>%
  filter(AcademicYear == "2020-2021")
gpa12_survey_19_20_21

gpa12_survey_19_20_21$num_visits_bin <-as.character(gpa12_survey_19_20_21$num_visits_bin)
gpa12_survey_19_20_21

#Regression
#gpa_12 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa12_survey_19_20_21)
#summary(gpa_12)

#Average GPA, 11th, (>2) visits to Camp RIO   (20-21)           GPA: 3.943
gpa12_survey_19_20_21_average_bin1 <- gpa12_survey_19_20_21%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/142)
gpa12_survey_19_20_21_average_bin1

#Average GPA, 11th, (1,2) visits to Camp RIO    (20-21)          GPA: 3.923
gpa12_survey_19_20_21_average_bin0 <- gpa12_survey_19_20_21%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/36)
gpa12_survey_19_20_21_average_bin0
#------------------------------------
#19-20
gpa12_survey_19<- gpa_survey_19_20_21 %>%
  filter(GradeLevelID == "12")%>%
  filter(AcademicYear == "2019-2020")
gpa12_survey_19

gpa12_survey_19$num_visits_bin <-as.character(gpa12_survey_19$num_visits_bin)
gpa12_survey_19

#Regression
#gpa_10_19 <- lm(CumulativeWeightedGPA ~ num_visits_bin, data=gpa10_survey_19)
#summary(gpa_10_19)

#Average GPA, 11th, (>2) visits to Camp RIO  (19-20)            GPA: 4.057
gpa12_survey_19_average_bin1 <- gpa12_survey_19%>%
  filter(num_visits_bin == 1)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/26)
gpa12_survey_19_average_bin1

#Average GPA, 11th, (1,2) visits to Camp RIO  (19-20)             GPA: 3.880
gpa12_survey_19_average_bin0 <- gpa12_survey_19%>%
  filter(num_visits_bin == 0)%>%
  mutate(gpa_average = sum(CumulativeWeightedGPA)/2)
gpa12_survey_19_average_bin0



```

ACT data
```{r}

act <- get_table(.table_name = "ACT", .database_name = "Dashboard", .schema = "dbo") %>%
  collect()
act

```
























